Visualizing demographic evolution using geographically inconsistent census data

### 2019-10-10 paper/study.tex
last paper review piccard 1.0 GEAN

\revision{1.12.(1-5)}{These results are meant to demonstrate the
  utility of the interface for understanding the evolutionary dynamics
  of urban neighbourhoods. They also show the face validity of the
  results generated by our novel network-based approach. While our
  method does not require geographic harmonisation, it does however
  require matching the variables over time. This is (currently) a
  painstaking, manual process, and the current version of the
  interface remains somewhat limited. For example, we selected a
  minimal set of aspects for each country, which are not necessarily
  similar to each other, or similar to studies in the current
  literature. Income moreover is slightly inaccurate, even though we
  did correct for official inflation. We grouped the original ranges
  into three larger ranges, but they do not match precisely. In
  addition, the US data labelled "2010" is not from the decennial
  census, but from the ACS 2006-2010, somewhat compromising the
  temporal stability of the data. Further, the regions selected do not
  correspond to any pre-defined regions (metros, cities, census
  areas), but to somewhat arbitrary regions defined around a location
  of interest, making direct comparisons harder.  Taken together,
  these factors would no doubt require caveats for a substantive study
  using the current prototype version of the interface. Nevertheless,
  in the present context we find that the interface results are
  well-aligned with several other studies, illustrating the robustness
  of our approach to imperfect conditions. }

@@

We consider a region that corresponds approximately to the administrative border
of the current city of Toronto, using all seven available aspects with equal
weights. While Chicago was fairly stable, Toronto is known to be a more dynamic
and diverse city, with significant and increasing immigrant
population~\citep{hulchanski2007three,Fong2011}, especially
Asian~\citep{Fong2003}. Toronto is also known for a stable and well defined
Jewish community~\citep{Harold2018, Fong2011}. Therefore, we expect \revision{1.14}{a}
combination of stable and dynamic regions on the results, with Place of Birth,
Home Language, and Religion identified as relevant aspects. The results are
summarised in Figure~\ref{fig:to}, considering eight clusters.

@@

\begin{figure}
    \centering 
    \includegraphics[width=0.75\linewidth]{toVol.pdf}
    \caption{Details for some regions of Toronto that were classified into 3 or
         more clusters over time.\label{fig:toVol}}
\end{figure}

@@

The overall overall response of the experts was positive,  mentioning that the
prototype allows them to analyse census data without the additional work of
obtaining and cleaning the data (A, B, E), and it allows the inclusion of
geographic visual analysis tools in their research process (D). It enables the
users to tell different stories about neighbourhoods/cities and their changes
(A), visualise the relationship between key urban variables over time (D),
offering a quick way to identify particular neighbourhoods that one may be
interested in studying more in depth around a particular issue or efficiently
understanding the context of an area (E).  Indeed, the experts identified
gentrification processes in Manhattan (B) and Dallas (E), reinforced a
hypothesis for occupational clustering (D), and highlighted how the method can
be used to compare neighbourhoods and cities (A). In summary, their view was that
the proposed methodology can be a viable alternative for the visual analytics of
evolving demographic data.

The interface was "easy to navigate" (B), but it was also considered
"overwhelming" (A), "intimidating" (E), and "tricky to interpret" (C), possible
side-effects of our effort to increase  representational accuracy, where we
avoided using simplified representation or labels. Identifying clusters by their
most relevant variables was welcome, but the overlap of information from
different clusters in the boxplot was "a bit confusing" (C) when colour was not
present. Further, most clusters can be sufficiently characterised using only the
most relevant aspect, but this is not generally true.

>>>>

These results are meant to demonstrate the utility of the interface for
understanding the evolutionary dynamics of urban neighbourhoods. They also show
the face validity of the results generated by our novel network-based approach.
While our method does not require geographic harmonisation, it does however
require matching the variables over time. This is (currently) a painstaking,
manual process, and the current version of the interface remains somewhat
limited. For example, we selected a minimal set of aspects for each country,
which are not necessarily similar to each other, or similar to studies in the
current literature. Income moreover is slightly inaccurate, even though we did
correct for official inflation. We grouped the original ranges into three larger
ranges, but they do not match precisely. In addition, the US data labelled
"2010" is not from the decennial census, but from the ACS 2006-2010, somewhat
compromising the temporal stability of the data. Further, the regions selected
do not correspond to any pre-defined regions (metros, cities, census areas), but
to somewhat arbitrary regions defined around a location of interest, making
direct comparisons harder.  Taken together, these factors would no doubt require
caveats for a substantive study using the current prototype version of the
interface. Nevertheless, in the present context we find that the interface
results are well-aligned with several other studies, illustrating the robustness
of our approach to imperfect conditions.

@@

We consider a region that corresponds approximately to the administrative border
of the current city of Toronto, using all seven available aspects with equal
weights. While Chicago was fairly stable, Toronto is known to be a more dynamic
and diverse city, with significant and increasing immigrant
population~\citep{hulchanski2007three,Fong2011}, especially
Asian~\citep{Fong2003}. Toronto is also known for a stable and well defined
Jewish community~\citep{Harold2018, Fong2011}. Therefore, we expect a
combination of stable and dynamic regions on the results, with Place of Birth,
Home Language, and Religion identified as relevant aspects. The results are
summarised in Figure~\ref{fig:to}, considering eight clusters.

@@

\begin{figure}
    \centering 
    \includegraphics[width=0.725\linewidth]{toVol.pdf}
    \caption{Details for some regions of Toronto that were classified into 3 or
         more clusters over time.\label{fig:toVol}}
\end{figure}

@@

The overall overall response of the experts was positive,  mentioning that the
prototype allows them to analyse census data without the additional work of
obtaining and cleaning the data (A, B, E), and it allows the inclusion of
geographic visual analysis tools in their research process (D). It enables the
users to tell different stories about neighbourhoods/cities and their changes
(A), visualise the relationship between key urban variables over time (D),
offering a quick way to identify particular neighbourhoods that one may be
interested in studying more in depth around a particular issue or efficiently
understanding the context of an area (E).  Indeed, the experts identified
gentrification processes in Manhattan (B) and Dallas (E), reinforced a
hypothesis for occupational clustering (D), and highlighted how the method can
be used to compare neighbourhoods and cities (A). In summary, their view was
that the proposed methodology can be a viable alternative for the visual
analytics of evolving demographic data. The interface was "easy to navigate"
(B), but it was also considered "overwhelming" (A), "intimidating" (E), and
"tricky to interpret" (C), possible side-effects of our effort to increase
representational accuracy, where we avoided using simplified representation or
labels. Identifying clusters by their most relevant variables was welcome, but
the overlap of information from different clusters in the boxplot was "a bit
confusing" (C) when colour was not present. Further, most clusters can be
sufficiently characterised using only the most relevant aspect, but this is not
generally true.
